---
title: "Getting Started with dimsprepr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with dimsprepr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

**dimsprepr** provides a complete, reproducible preprocessing pipeline for
direct-injection mass spectrometry (DIMS) metabolomics data.  It covers:
missing-value filtering, imputation, signal-drift and batch correction,
normalization, transformation, scaling, and quality filtering — all
orchestrated through a single function call.

## Quick Example

```{r setup}
library(dimsprepr)

# Load the bundled example dataset
data(dims_data)
data(dims_metadata)

dim(dims_data)
table(dims_metadata$Group)
```

### Run the full pipeline

The `run_DIpreprocess()` function executes all preprocessing steps in a
fixed, validated order.  Here we skip drift correction (since this is a
small simulated dataset without a realistic QC injection pattern):
```{r pipeline}
result <- run_DIpreprocess(
  x                = dims_data,
  metadata         = dims_metadata,
  normalize_method = "median",
  transform_method = "log10",
  correct_drift    = FALSE,
  rsd_threshold    = 0.5,
  variance_percentile = 5,
  verbose          = FALSE
)

result
```

The result is a named list with intermediate and final data matrices:

```{r inspect}
# Final preprocessed matrix (auto-scaled for general analysis)
dim(result$data_nonpls)

# Step-by-step dimension tracking
result$dimensions
```

## Using Individual Functions

Each preprocessing step is also available as a standalone function for
maximum flexibility.

### 1. Missing-Value Filtering

```{r filtermissing}
filt <- run_filtermissing(
  dims_data,
  dims_metadata,
  threshold = 0.3,
  verbose   = FALSE
)
filt
```

### 2. Imputation

```{r impute}
imp <- run_mvimpute(filt$data, method = 0.2, verbose = FALSE)
imp
```

### 3. Normalization

```{r normalize}
norm <- run_normalize(imp$data, dims_metadata, method = "median", verbose = FALSE)
```

### 4. Transformation

```{r transform}
trans <- run_transform(norm$data, method = "log10", verbose = FALSE)
```

### 5. Scaling

```{r scale}
scaled <- run_scale(trans$data, method = "auto", verbose = FALSE)
scaled
```

## Cartesian-Product Pipelines

Supply vectors to `normalize_method` and/or `transform_method` to
compute all combinations automatically.  Expensive upstream steps
(filtering, imputation, drift correction) run only once:

```{r multi, eval=FALSE}
multi <- run_DIpreprocess(
  x                = dims_data,
  metadata         = dims_metadata,
  normalize_method = c("sum", "median"),
  transform_method = c("log10", "log2"),
  correct_drift    = FALSE,
  verbose          = FALSE
)
names(multi)
# [1] "sum_log10"    "sum_log2"     "median_log10" "median_log2"
```

## Session Info

```{r session}
sessionInfo()
```
